---
title: "Histogram of the number of site per protein"
author: "Marie Locard-Paulet"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r Packages, echo=FALSE, message=FALSE, warning=FALSE}
require(ggplot2)
require(knitr)
```

```{r}
load("RData/12_PhosphoTableWithProteinWaring.RData")
load(file = "RData/03_psiteInfoPhosphoSitePlus.RData")
```

# Histogram of the number of detected phosphorylation sites per proteins

```{r}
for (i in seq_along(psiteInfo)) {
  vec <- psiteInfo[[i]]
  vec$Accession <- strsplit(vec$phosphoSites, "_", fixed = T)[[1]][1]
  psiteInfo[[i]] <- vec
}
accessions <- sapply(psiteInfo, function(x) {
  x$Accession
})
monosites <- sapply(psiteInfo, function(x) {
  x$MonoSites
})
accessions <- sapply(seq_along(accessions), function(x) {
  rep(accessions[x], length(monosites[[x]]))
})
tabcountprot <- data.frame("Accession" = unlist(accessions), "Monosite" = unlist(monosites))
tabcountprot <- tabcountprot[!duplicated(tabcountprot),]
gtab <- data.frame("Protein"= names(table(as.character(tabcountprot$Accession))), "pSiteNumber"= as.numeric(as.character(table(tabcountprot$Accession))))
ggplot(gtab, aes(x = pSiteNumber)) + geom_histogram(bins = 100) + theme_bw() + coord_flip()
```

I reduce the y-axis by removing the proteins with more than 30 sites.

Table of removed proteins:

`r kable(gtab[gtab[,2]>30,])`

```{r}
g <- ggplot(gtab, aes(x = pSiteNumber)) + geom_histogram(col = "black", fill = "grey70") + theme_minimal() + xlim(0,30) + xlab("Number of phosphorylation site nper protein") + ylab("Frequency") + geom_hline(yintercept = 0) # + coord_flip()
print(g)
pdf("Figures/HistPsites.pdf", useDingbats=FALSE, 8, 8)
print(g)
dev.off()
```

# I do the same for regulated proteins

```{r}
accessions <- sapply(psiteInfo[names(psiteInfo) %in% export$psiteID[export$Regulation == "TRUE"]], function(x) {
  x$Accession
})
monosites <- sapply(psiteInfo[names(psiteInfo) %in% export$psiteID[export$Regulation == "TRUE"]], function(x) {
  x$MonoSites
})
accessions <- sapply(seq_along(accessions), function(x) {
  rep(accessions[x], length(monosites[[x]]))
})
tabcountprot <- data.frame("Accession" = unlist(accessions), "Monosite" = unlist(monosites))
tabcountprot <- tabcountprot[!duplicated(tabcountprot),]
gtab <- data.frame("Protein"= names(table(as.character(tabcountprot$Accession))), "pSiteNumber"= as.numeric(as.character(table(tabcountprot$Accession))))

ggplot(gtab, aes(x = pSiteNumber)) + geom_histogram(bins = 100) + theme_bw() + coord_flip()
ggplot(gtab, aes(x = pSiteNumber)) + geom_histogram() + theme_bw() + coord_flip()
```

I reduce the y-axis:

```{r}
g <- ggplot(gtab, aes(x = pSiteNumber)) + geom_histogram(col = "black", fill = "grey70") + theme_minimal() + xlab("Number of phosphorylation site nper protein") + ylab("Frequency") + geom_hline(yintercept = 0) # + coord_flip()
print(g)
pdf("Figures/HistPsitesReg.pdf", useDingbats=FALSE, 8, 8)
print(g)
dev.off()
```

# Relation between length of the protein and our ability to detect phosphorylation sites

I also load the Uniprot database for mouse (only reviewed proteins, downloaded the 20th of November 2018) in order to check the depth of our analysis.
```{r}
# First, I have removed the characters "'" and "#".
up <- read.table(file = "UniprotReviewedMouse_20190920/uniprot-filtered-organismMusmusculusReviewed.tab", sep = "\t", header = T)
```

Map the protein length to the table of proteins and plot the correlation.

**I remove the protein Titin that is really long compared to the others: 35213aa.**

```{r}
accessions <- sapply(psiteInfo, function(x) {
  x$Accession
})
monosites <- sapply(psiteInfo, function(x) {
  x$MonoSites
})
accessions <- sapply(seq_along(accessions), function(x) {
  rep(accessions[x], length(monosites[[x]]))
})
tabcountprot <- data.frame("Accession" = unlist(accessions), "Monosite" = unlist(monosites))
tabcountprot <- tabcountprot[!duplicated(tabcountprot),]
gtab <- data.frame("Protein"= names(table(as.character(tabcountprot$Accession))), "pSiteNumber"= as.numeric(as.character(table(tabcountprot$Accession))))
gtab$length <- up$Length[match(as.character(gtab$Protein), up$Entry)]
gtab$length <- as.numeric(as.character(gtab$length))
gtab <- gtab[gtab$length<30000,]
gtab <- gtab[!is.na(gtab$pSiteNumber),]
# Relationship between length and "detectability":
ggplot(gtab, aes(x = pSiteNumber, y = length, group = pSiteNumber)) + geom_violin(alpha = 0.6 , fill = "grey70")  + theme_minimal() + geom_dotplot(binaxis= "y",stackdir ="centerwhole", dotsize = 0.1, stackratio = 0.1, alpha = 0.6) + xlim(0,30)

ggplot(gtab, aes(x = pSiteNumber, y = length, group = pSiteNumber))   + geom_jitter(alpha = 0.5, size = 0.6, position = position_jitter(height = .1, width = .2))  + theme_minimal() + xlim(0,30) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + xlab("Number of phosphorylation site identified per protein")

pdf("Figures/NumPsiteVsLength.pdf", useDingbats=FALSE, 8, 8)
ggplot(gtab, aes(x = pSiteNumber, y = length, group = pSiteNumber))   + geom_jitter(alpha = 0.5, size = 0.6, position = position_jitter(height = .1, width = .2))  + theme_minimal() + xlim(0,30) + geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + xlab("Number of phosphorylation site identified per protein")
dev.off()
```

I save the table for mapping the number of site per protein in other graphical representations:

```{r}
save(gtab, file = "RData/NumSitePerProt.RData")
```


```{r SessionInfo, echo=FALSE}
sessionInfo()
``` 
