---
title: "TMT analysis of OT-1 T-cells upon TCR stimulation"
author: "Marie Locard-Paulet"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---


```{r Packages, echo=FALSE, message=FALSE, warning=FALSE}
library(gplots)
library(knitr)
library(reshape2)
library(ggplot2)
library(dbscan)
library(missMDA)

knitr::opts_chunk$set(echo = FALSE)

colfuncClusters <- colorRampPalette(c("#9E0142", "#D53E4F", "#F46D43", "#FDAE61", "#ABDDA4", "#66C2A5", "#3288BD", "#5E4FA2", "darkblue", "black"))
colourClus <- colfuncClusters(12)
```


```{r}
load(file = "RData/11_TableClusters.RData")
tab <- export
source("RFunctions/PCAPlots.R")
load(file = "RData/03_psiteInfoPhosphoSitePlus.RData")
```

I use as input the fold changes upon TCR activation.

# PCA plots

```{r}
mat <- tab[,grepl(".FC", names(tab), fixed = T)]
row.names(mat) <- tab$psiteID
# Keep only the regulated sites:
mat <- mat[row.names(mat) %in% tab$psiteID[tab$Regulation == "TRUE"],]

# # Imputation of missing data:
# nb <- estim_ncpPCA(mat, scale = F)
# comp <- imputePCA(mat, ncp = 2, scale = F)
# 
# pca <- prcomp(comp$completeObs, scale = F, center = F)

# I don't have missing values so:
pca <- prcomp(mat, scale = F, center = F)
plotSDev(pca)
plotGenes(pca)

gtab <- data.frame("x" = pca$x[,1], "y" = pca$x[,2], "ID" = row.names(pca$x))
gtab$Cluster <- tab$ClusterMerged[match(gtab$ID, tab$psiteID)]
gtab$GeneID <- sapply(gtab$ID, function(x) {
    psiteInfo[[which(names(psiteInfo) == x)]]$GeneID
  })
gtab$labels <- gtab$GeneID
gtab$labels[!grepl("Cd", gtab$labels)] <- ""
ggplot(gtab, aes(x = x, y = y, col = factor(Cluster), label = labels)) + geom_point(alpha = 0.6, size = 3) + scale_color_manual(values = colourClus) + theme_bw() + labs(title = "PCA of the regulated sites") # + geom_text(size = 4)
```
================================================================================

# Functional analysis

I find from PhosphoSitePlus the putative kinases responsible for the phosphorylation sites regulations.

```{r}
load(file = "RData/DBMapping/PsitePlus.RData")
temp <- sapply(export$psiteID, function(x) {
    psiteInfo[[which(names(psiteInfo) == x)]]$PhosphoSitePlusAll[1]
  })
export$PsitePlusIDMultiSpecies <- temp
names(tabK)[names(tabK) == "ALL_SITEGPID"] <- "PsitePlusIDMultiSpecies"
tabK <- tabK[tabK$KIN_ORGANISM %in% c("human", "mouse", "rat"),]
tabK <- tabK[order(tabK$KIN_ORGANISM),]
mer <- merge(export, tabK, by = "PsitePlusIDMultiSpecies", all = T)
mer <- mer[!is.na(mer$psiteID),]

temp <- sapply(gtab$ID, function(x) {
  paste(unique(mer$KINASE[as.character(mer$psiteID) == x]), collapse = "|")
})
gtab$Kinase <- temp
```

I plot the PCAs with tthe names of all the kinases that are found regulating the sites:

```{r}
kinases <- unique(
  unlist(
    sapply(gtab$ID, function(x) {
      unique(mer$KINASE[as.character(mer$psiteID) == x])
    })
  )
)
kinases <- sort(kinases)

pdf("Figures/PCAKinases.pdf", useDingbats=FALSE, 9, 8)
for (el in kinases) {
  gtab$labels <- gtab$GeneID
  gtab$labels[!grepl(el, gtab$Kinase)] <- ""
  g <- ggplot(gtab, aes(x = x, y = y, col = factor(Cluster), label = labels)) + geom_point(alpha = 0.3, size = 3) + scale_color_manual(values = colourClus) + theme_bw() + labs(title = "PCA of the regulated sites", subtitle = el)  + geom_text(size = 4, fontface = "bold")
  print(g)
}
dev.off()
```

I plot the plots of the PCA with the kinase substrates in the document "Figures/PCAKinases.pdf".

```{r}
temp <- gtab[,c(3,7)]
temp <- temp[temp$Kinase != "NA",]
temp <- temp[!duplicated(temp),]
write.table(temp, "SupTables/PhosphoSitePlusMapping.txt", sep = "\t", row.names = F, quote = F)
```

================================================================================

```{r SessionInfo, echo=FALSE}
sessionInfo()
``` 