---
title: "TMT analysis of OT-1 T-cells upon TCR stimulation"
author: "Marie Locard-Paulet"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---


```{r Packages, echo=FALSE, message=FALSE, warning=FALSE}
library(gplots)
library(knitr)
library(reshape2)
library(ggplot2)
library(dbscan)
library(missMDA)

knitr::opts_chunk$set(echo = FALSE)

colfuncClusters <- colorRampPalette(c("#9E0142", "#D53E4F", "#F46D43", "#FDAE61", "#ABDDA4", "#66C2A5", "#3288BD", "#5E4FA2", "darkblue"))
colourClus <- colfuncClusters(13)
```


```{r}
load(file = "RData/11_TableClusters.RData")
tab <- export
load(file = "RData/03_psiteInfoPhosphoSitePlus.RData")
tsne <- read.table("tsne/table_all_phospho_loca_tSNENew2018_NewClusters.csv", sep = ";", header = T)
```

I use as input the fold changes upon TCR activation.

# tsne plots

```{r}
ggplot(tsne, aes(x = V1, y = V2, col = factor(NewClusters))) + geom_point(alpha = 0.6, size = 3) + scale_color_manual(values = colourClus) + theme_bw() + labs(title = "tsne of the regulated sites") 
```
================================================================================

# Functional analysis

I find from PhosphoSitePlus the putative kinases responsible for the phosphorylation sites regulations.

```{r}
load(file = "RData/DBMapping/PsitePlus.RData")
tempM <- sapply(export$psiteID, function(x) {
    psiteInfo[[which(names(psiteInfo) == x)]]$PhosphoSitePlusMouse[1]
  })
tempA <- sapply(export$psiteID, function(x) {
    psiteInfo[[which(names(psiteInfo) == x)]]$PhosphoSitePlusNotMouse[1]
  })
tempM[is.na(tempM)] <- tempA[is.na(tempM)]
export$PsitePlusIDMultiSpecies <- tempM
tabK <- ltabK$KinaseSubstrate
names(tabK)[names(tabK) == "SITE_GRP_ID"] <- "PsitePlusIDMultiSpecies"
tabK <- tabK[tabK$KIN_ORGANISM %in% c("human", "mouse", "rat"),]
tabK <- tabK[order(tabK$KIN_ORGANISM),]
mer <- merge(export, tabK, by = "PsitePlusIDMultiSpecies", all = T)
mer <- mer[!is.na(mer$psiteID),]

temp <- sapply(tsne$kineticMean.psiteID, function(x) {
  paste(unique(mer$KINASE[as.character(mer$GeneID) == x]), collapse = "|")
})
tsne$Kinase <- temp
```

I plot the PCAs with the names of all the kinases that are found regulating the sites:

```{r}
kinases <- unique(
  unlist(
    sapply(tsne$kineticMean.psiteID, function(x) {
      unique(mer$KINASE[as.character(mer$GeneID) == x])
    })
  )
)
kinases <- sort(kinases)
tsne$Kinase <- as.character(tsne$Kinase)

pdf("Figures/tsneKinases.pdf", useDingbats=FALSE, 9, 8)
for (el in kinases) {
  tsne$labels <- as.character(tsne$kineticMean.psiteID)
  tsne$labels[!grepl(el, tsne$Kinase)] <- ""
  g <- ggplot(tsne, aes(x = V1, y = V2, col = factor(NewClusters), label = labels)) + geom_point(alpha = 0.3, size = 3) + scale_color_manual(values = colourClus) + theme_bw() + labs(title = "tsne of the regulated sites", subtitle = el)  + geom_text(size = 4, fontface = "bold")
  print(g)
}
dev.off()
```

I plot the plots of the tsne with the kinase substrates in the document "Figures/tsneKinases.pdf".

```{r, eval = F}
temp <- tsne[,c(3,7)]
temp <- temp[temp$Kinase != "NA",]
temp <- temp[!duplicated(temp),]
write.table(temp, "SupTables/PhosphoSitePlusMapping.txt", sep = "\t", row.names = F, quote = F)
```

```{r}
pdf("Figures/tsne.pdf", useDingbats=FALSE, 9, 8)
g <- ggplot(tsne, aes(x = V1, y = V2, col = factor(NewClusters))) + 
  geom_point(alpha = 0.6, size = 3) + 
  scale_color_manual(values = colourClus) + 
  theme_bw() + 
  labs(title = "tsne of the regulated sites")  
print(g)
dev.off()
```

================================================================================

```{r SessionInfo, echo=FALSE}
sessionInfo()
``` 